Block bootstrap consistency\\under weak assumptions

### 2015-12-31 Paper.tex
[#7] Address conditioning issues raised by referee
This commit tries to expalain some of the conditioning issues that
bother one of the referees.

First, the relevant probability infrastructure.  Let
$(S, \Omega, \pr)$ be a probability space and define the sequence of
sub-sigma-fields $\Omega_n \subset \Omega$. Assume that each
$X_{nt}$, $V_{nt}$, $M_{nj}$, and $u_{nj}$ is $\Omega_n$-measurable,
with $u_{nj}$ the $\operatorname{uniform}(1,\dots,n)$ random variables
that designate the start period of each bootstrap ``block.'' Also
define the \sigmafield\ generated by the stationary bootstrap's block
lengths alone,
\begin{equation}
  \Ms = \sigma(J_{n}, M_{n1},\dots,M_{nJ_n})
\end{equation}
and the conditional probability $\Pm(\cdot) = \pr[\cdot \mid \Ms]$.
(And define $\Pm^*(\cdot) = \pr^*[\cdot \mid \Ms]$, $\Em(\cdot) =
\E^*(\cdot \mid \Ms)$, etc.) An important property is
that $\Ms$ is \emph{independent} of the $X_{nt}$, so we can treat any
$M_{nj}$ and $J_n$ terms as constants within $\Em(\cdot)$ and
$\Pm[\cdot]$ and integrate over the unconditional distributions of the
$X_{nt}$. This property is especially important because it allows us
to use maximal inequalities and other moment inequalities for
mixingale arrays without modification to construct almost sure bounds on the
conditional moments $\Em$.

@@

To prove~\eqref{eq:19}, it suffices to show that
\begin{equation}\label{eq:21}
  \Em \sum_{j=1}^{J_n} \Em^{*}\big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2} > \epsilon\}\big) \to^p 0.
\end{equation}
Lemma~\ref{bootstrapped-zsq-uniform-integrability-lemma} implies that there exists a finite,
monotone function $B(\cdot)$ such that $B(x) \to 0$ as $x \to \infty$
and
\begin{equation*}
  \Em( (n Z_{nj}^{*2} / M_{nj}) \ind\{n Z_n^{*2} / M_{nj} > \epsilon n / M_{nj}\}) \\
  \leq B(\epsilon n /M_{nj})
\end{equation*}
almost surely for large enough $n$ and all $j$ and $\tau$.  Consequently,
\begin{align*}
    \Em \sum_{j=1}^{J_n} \Em^{*}( Z_{nj}^{*2} \ind\{Z_{nj}^{*2} > \epsilon\})
    & \leq \sum_{j=1}^{J_n} ( M_{nj} / n )\, B(\epsilon n /M_{nj}) \\
    & \leq \max_{j = 1,\dots,J_n} B(\epsilon n / M_{nj}) \sum_{i=1}^{J_n} M_{ni} / n \\
    & = B(\epsilon n / \max_{j = 1,\dots,J_n} M_{nj}) \\
    & \to^p 0 \text{\ as\ } n \to \infty,
\end{align*}
where the equality holds by monotonicity of $B$ and convergence in
probability holds by Lemma~\ref{block-length-convergence-lemma},
completing the proof of~\eqref{eq:19}.
The Dominated Convergence Theorem then ensures that
\begin{equation}\label{eq:22}
  \pr^{*} \Big[\sum_{j=1}^{J_n}
  Z_{nj}^{*} / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0.
\end{equation}
follows from~(\ref{eq:18}). (Also see
Lemma~\ref{equivalent-convergence-from-lie-lemma}.)

@@

@@

\begin{proof}
  See supplemental appendix for the proof of~\eqref{eq:34}, which
  follows \citet[Lemma 6.5]{Mcl:75b} and \citet[Lemma 3.5]{Mcl:77}
  almost exactly and is also presented as Theorem~16.13 in
  \citet{Dav:94}.
\end{proof}

>>>>

First, the relevant probability infrastructure.  Let
$(S, \Omega, \pr)$ be a probability space and define the sequence of
sub-sigma-fields $\Omega_n \subset \Omega$. Assume that each
$X_{nt}$, $V_{nt}$, $M_{nj}$, and $u_{nj}$ is $\Omega_n$-measurable,
with $u_{nj}$ the $\operatorname{uniform}(1,\dots,n)$ random variables
that designate the start period of each bootstrap ``block.'' Also
define the \sigmafield\ generated by the stationary bootstrap's block
lengths alone,
\begin{equation}
  \Ms = \sigma(J_{n}, M_{n1},\dots,M_{nJ_n})
\end{equation}
and the conditional probability $\Pm(\cdot) = \pr[\cdot \mid \Ms]$.
(And define $\Pm^*(\cdot) = \pr^*[\cdot \mid \Ms]$, $\Em(\cdot) =
\E^*(\cdot \mid \Ms)$, etc.) An important property is
that $\Ms$ is \emph{independent} of the $X_{nt}$'s, $V_{nt}$'s, and
$u_{nj}$'s, so we can treat any
$M_{nj}$ and $J_n$ terms as constants within $\Em(\cdot)$ and
$\Pm[\cdot]$ and integrate over the unconditional distributions of the
other random variables. This property is especially important because it allows us
to use maximal inequalities and other moment inequalities for
mixingale arrays with only small modifications to construct almost sure bounds on the
conditional moments $\Em$.

@@

To prove~\eqref{eq:19}, it suffices to show that
\begin{equation}\label{eq:21}
  \Em \sum_{j=1}^{J_n} \Em^{*}\big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2} > \epsilon\}\big) \to^p 0.
\end{equation}
For any $j$ and $m$ and for large enough $n$,
\begin{align}
\Em \Em^{*}\big((Z_n(u_{nj}, m)^2n/m) &\ind\{Z_n(u_{nj}, m)^2n/m > \epsilon n/m\}\big) \notag \\
&= \E \big((Z_n^{*}(0, m)^2n/m) \ind\{Z_n^{*}(0, m)^2n/m > \epsilon n/m\}\big) \notag \\
&\leq B(\epsilon n/m)\label{eq:5}
\end{align}
where $B$ is some finite monotone function such that $B(x) \to 0$ as
$x \to \infty$; the existence of a function $B$ that satisfies these
conditions (and does not depend on $n$ or $m$) is a consequence of
Lemma~\ref{bootstrapped-zsq-uniform-integrability-lemma},
Equation~\eqref{eq:30}. Since $J_n$ and $M_{nj}$ are both $\Ms$-measureable
and independent of the $u_{nj}$'s and $X_{nt}$'s,~\eqref{eq:5} implies that
\begin{align*}
\Em &\sum_{j=1}^{J_n} \Em^{*}\big(Z_{nj}^{*2} \ind\{Z_{nj}^{*2} > \epsilon\}\big) \\
& = \sum_{j=1}^{J_n} (M_{nj}/n) \Em \Em^{*}
\big((Z_n(u_{nj}, M_{nj})^2 n/M_{nj})
\ind\{Z_n(u_{nj}, M_{nj})^2n/M_{nj} > \epsilon n/M_{nj}\}\big) \\
& \leq \sum_{j=1}^{J_n} (M_{nj}/n) \, B(\epsilon n/M_{nj}) \\
& \leq \max_{j = 1,\dots,J_n} B(\epsilon n / M_{nj}) \sum_{i=1}^{J_n} M_{ni} / n \\
& = B(\epsilon n / \max_{j = 1,\dots,J_n} M_{nj}) \\
& \to^p 0 \text{\ as\ } n \to \infty.
\end{align*}
The first equality follows from simple algerbra and the definition
of the $u_{nj}$'s. The first inequality is a
consequence of~\eqref{eq:5}.  The second equality holds by
monotonicity of $B$. And convergence in probability holds by
Lemma~\ref{block-length-convergence-lemma}, completing the proof
of~\eqref{eq:19}. The Dominated Convergence Theorem and~(\ref{eq:18})
then ensure that
\begin{equation}\label{eq:22}
  \pr^{*} \Big[\sum_{j=1}^{J_n}
  Z_{nj}^{*} / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0.
\end{equation}
(Also see
Lemma~\ref{equivalent-convergence-from-lie-lemma}.)

@@

Note that these results apply immediately to the conditional expectation
$\Em$ because the block lengths $M_{jn}$ do not appear in~\eqref{eq:32}
and~\eqref{eq:33} and are independent of all of the random variables
used for these bounds.

@@

For proof, see supplemental appendix for the proof of~\eqref{eq:34}, which
follows \citet[Lemma 6.5]{Mcl:75b} and \citet[Lemma 3.5]{Mcl:77}
almost exactly and is also presented as Theorem~16.13 in
\citet{Dav:94}. The same comment that follows
Lemma~\ref{dejong-restatement-lemma} applies here as well: the bounds
apply equally well to $\Em$.
